### 2020-01-03 yfcc100m/paper/vldb2020/intro.tex

\section{Introduction}
\label{intro}

Visual computing workloads performing analytics on
video or image data, either off-line or streaming,
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning techniques to
extract information out of the visual data which can subsequently be used
for informed decision making.
The insights this information can provide depend on the
application: a retail vendor might be interested in the amount of time
shoppers spend in front of a specific product, while a medical expert might
want to see the effect of a specific treatment on the size of a tumor.

Despite this rich and varied usage environment, there has been very little
research on the management of visual data.
Most of the current storage solutions are
an ad-hoc collection of tools combined with custom scripts to tie them
together, unique not only to a specific discipline but often to individual
researchers. For example, consider an ML developer constructing a pipeline
for extracting brain tumor information from existing brain images in a
classic medical imaging use case. This requires assigning consistent
identifiers for the scans and adding their metadata in
some form of relational or key-value database. If the queries require
search over some patient information, then patients have to be associated
with their brain scans. Finally, if the ML pipeline needs images that
are of a size different than the stored ones, there is additional compute
diverted towards preprocessing after the potentially larger images are
fetched. All these steps require investigation of different software
solutions that provide various functionalities that can then be stitched
together with a script for this specific use case.
Moreover, if the pipeline identifies
new metadata to be added for the tumor images, most databases make it
hard to evolve the schema on the fly.
Not only do these ad hoc solutions make replicating experiments
difficult, they do not scale well to real-world applications.
Addressing the storage and retrieval of visual data necessitates a complete
overhaul of the storage architecture,
preferably using emerging breakthroughs in
heterogeneous memory and storage for efficiency.

We present the Visual Data Management System (VDMS)\cite{darkside},
an Open Source project designed to enable efficient access of visual data.
Since visual data often contains
rich metadata (such as objects, locations, and time), VDMS stores this
information in a high performance graph database. Using this metadata, VDMS
can quickly identify which data is relevant to a given query.
Additionally, VDMS uses a custom library to store and retrieve visual data,
which provides an interface for machine friendly formats as well as
traditional formats. These new formats are designed to support applications
that are often interested in specific areas of images or videos,
particularly when the individual object is large.

While there are a number of big-data frameworks~\cite{spark, hadoop}, systems
that can be used to store metadata~\cite{memsql, vertica}, and systems that
manipulate a specific category of visual data~\cite{scidb, rasdaman}, VDMS can
be distinguished from them on the following aspects:

\begin{itemize}
\item {\em Design for analytics and machine learning}: by targeting
visual data for use cases that require manipulation
of visual information and associated metadata,
\item {\em Ease-of-use}: By defining a common API that allows applications to
combine their complex metadata searches with operations on resulting visual
data, and together with full support for feature vectors, VDMS goes beyond the
traditional SQL or OpenCV level interfaces that do one or the other. Given our
focus on enabling machine learning, we also provide a client API in Python.
\item {\em Performance}: We show how a unified system such as VDMS can
outperform an ad-hoc system constructed with well-known discrete components.
Because of the capabilities we have built into VDMS, it handles complex
queries significantly better than the ad-hoc system without compromising the
performance of simple queries.
\end{itemize}