### 2020-01-15 yfcc100m/paper/vldb2020/conclusion.tex
Add more figures and other minor additions

\section{Conclusion}

We introduced the Visual Data Management System,
designed to enable efficient access of visual data.
We presented our rich, JSON-based API, designed to simplify visual data access
for data scientists and analytics pipelines.
We compared the performance of our system to a
baseline of a combination of widely available and used systems.
Our findings showed that VDMS efficiently deals with complex queries,
providing a performance improvement
of up to 2x in the examined medical data search use-case.
Furthermore, VDMS requires significantly fewer
lines of code to execute complex queries in complex visual pipelines.
We intend to continue with the evaluation of
VDMS performance different use cases, and to identify more
opportunities to optimize and simplify the access of visual data
for machine learning and analytics applications.