### 2020-02-03 yfcc100m/paper/vldb2020/eval.tex

\section{Evaluation}
\label{eval}

\subsection{YFCC100M Dataset}
\label{dataset}

The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a large collection of 100 million public Flickr media objects created to provide free, sharable multimedia data for research. This dataset contains approximately 99.2 million images and 0.8 million videos with metadata characterized by 25 fields such as the identifier, user, date the media was taken/uploaded, location in longitude/latitude coordinates, device the media was captured, URL to download the media object, and Creative Commons license information.  The YFCC100M dataset also used a deep learning approach to generate autotags which is a set of comma-separated concepts such as people, scenery, objects, and animals and confidence scores generated from 1,570 trained Caffe classifiers ~\cite{Thomee_2016}.
We have also used feature vectors generated for every image and first frame
of every video \cite{features} to implement similarity search.

\subsection{Experimental Setup}

For all our experiments, we use 2 servers, one hosting VDMS server and
another hosting MySQL server. Both servers have a dual-socket
Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with multithreading enabeled,
for a total of 112 logical cores per server.
The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
The servers and client are connected through a 1GB wired link.

As a baseline, we implemented a similar database system comprised of a
combination of widely available and out-of-the-box components: MySQL Server 5.7,
Apache Web Server 2.4.18, and OpenCV 3.3.
Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (which wants to retrieve metadata and
imaged) with VDMS (left) and the baseline (right).
It is worth noting that the images are stored in a shared repository
(ext4 filesystem on a RAID 6 configuration of 16TB) that both Apache WebServer
and VDMS have direct access. In the case of MySQL, metadata is stored in an
attached SSD disk. Even if VDMS has native support for Optane Persistent Memory,
we do not use it in this experiment because of fairness of comparison w.r.t
MySQL. The benefits of Persistent Memory on metadata operations is left
for another paper, and outside the scope of this evaluation.
For this experiment then, we simply use a similar attached SSD disk to store
metadata (even if VDMS Graph store is designed for PM, it can deliver good
performance when using SSDs directly).

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/comparison_system}
\caption{Comparison Systems}
\label{fig:systems}
\end{figure*}

We built VDMS and MySQL databases using the YFCC100M dataset with incremental database sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M. For each database size, we created a PMGD graph from the YFCC100m metadata, the YFCC100m autotags associated with each metadata identifier, and a list of 1,570 autotags using 100 concurrent VDMS Python clients.  With simple queries, we insert 1,570 nodes for the autotags and insert nodes for the YFCC media objects with the appropriate metadata.  Adding the autotag confidence scores is a recursive process and requires more complex queries than adding the previous nodes. For each metadata identifier in autotags, the query must find the associated metadata node and tag node.  A connection between the two nodes is created with the confidence score as a property. The number of metadata nodes are dependent on the database size and the connections are responsible for 90\% of the elements in each database as shown in Table~\ref{table:vdmsnodes}. It is important to note that the metadata identifier, autotags, and longitude/latitude coordinates are set as indexes of the database to allow faster retrieval of the metadata.

\begin{table}[h]
\caption{Nodes in VDMS database}
\centering
\begin{tabular}{c c c c}
\hline\hline
Database Size & Connections & Images & Autotags\\
\hline
100k & 848,432      & 100,000     & 1,570\\
500k & 4,249,500    & 500,000     & 1,570\\
1M   & 8,503,045    & 1,000,000   & 1,570\\
5M   & 42,505,478   & 5,000,000   & 1,570\\
10M  & 85,040,404   & 10,000,000  & 1,570\\
50M  & 425,162,070  & 50,000,000  & 1,570\\
100M & 895,572,430  & 99,205,984  & 1,570\\
\hline
\end{tabular}
\label{table:vdmsnodes}
\end{table}

Each MySQL database is created in a similar manner but the data is represented as three tables: taglist, metadata, and autotags.  By default, MySQL uses inifinite number of threads inside InnoDB when processing requests using four threads per IO read/write.  In some cases, when creating large databases data locks may occur to protect the data from concurrent updates.  To minimize locking in our runs, we increased the InnoDB buffer pool size to increase the amount of memory allocated to in-memory data structures ~\cite{mysql,mysql_blog}. Using a Python client and simple queries, the taglist table is read from the list of tags with an auto-incremented tagid as an index and the metadata table is read from the YFCC100m metadata using the identifier as an index. The autotags table contains the generated autotags and confidence scores for metadata entries of the metadata table. To generate the table, we split the autotags data for each database by the metadata identifier and autotag into new files. The new files are read into the respective autotags table with the metadata identifier and tagid as indexes.

\begin{table}[h]
\caption{Rows in MySQL database}
\centering
\begin{tabular}{c c c c}
\hline\hline
 & \multicolumn{3}{c}{Table}\\
\cline{2-4}
Database Size & Autotags & MetaData & Tag List\\
\hline
100k & 848,912     & 100,000    & 1,570\\
500k & 4,241,200   & 498,707    & 1,570\\
1M   & 8,508,380   & 1,000,000  & 1,570\\
5M   & 42,425,905  & 4,987,379  & 1,570\\
10M  & 85,095,265  & 10,000,000 & 1,570\\
50M  & 425,446,208 & 50,000,000 & 1,570\\
100M & 896,002,496 & 99,206,564 & 1,570\\
\hline
\end{tabular}
\label{table:mysqltables}
\end{table}

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/db_time_size}
\caption{Time (in hours) to build and size (in GB) of MySQL and VDMS databases}
\label{fig:db_time_size}
\end{figure}

The VDMS and MySQL databases have comparable number of elements as shown in Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables} but VDMS outperforms MySQL in build speed.  Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases faster than MySQL as the database size grows.  Key difference in the build times are contributed to the low-level implementation of how MySQL reads and stores data from the files and the optimizations (increased InnoDB pool size, etc.) needed to handle large datasets such as YFCC100m.  On average, it took MySQL 3.72x hours longer to build each database than VDMS. For example, to build the 100M database, MySQL took 263 hours while VDMS only needed 72.5 hours.  This is a difference of over 7 days of processing data with VDMS missing less than 0.001\% images and 0.048\% connections from the entire dataset.  Alternatively, VDMS requires more storage, shown in Figure~\ref{fig:db_time_size}, to store information about each node/connection.  This may become a factor if storage is a limitation.  For our runs, VDMS required 30-41\% more storage than MySQL.

\subsection{Images + Metadata}

To evaluate the access to metadata and images, we use the following four queries:
\begin{enumerate}
\item {\bf {\em 1tag}}: Find metadata/images of alligator.
\item {\bf {\em 1tag\_resize}}: Find metadata/images of alligator and resize to 224x224.
\item {\bf {\em 1tag\_resize\_geo}}: Find metadata/images of alligator, resize to 224x224, and in a particular geolocation (20 degrees radius).
\item {\bf {\em 2tag\_resize}}: Find metadata/images of BOTH alligator and lake and resize to 224x224.
\end{enumerate}

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/concurrency_vdms}
\caption{Concurrency Analysis - VDMS}
\label{fig:concurrency_vdms}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/queries_throughput_32}
\caption{Queries - Throughput - 32 Clients}
\label{fig:q_throughput_32}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/queries_throughput_56}
\caption{Queries - Throughput - 56 Clients}
\label{fig:q_throughput_56}
\end{figure}

An image and metadata search on the YFCC100M databases are very extensive
and would benefit from multithreading to process queries simultaneously.
Figure~\ref{fig:concurrency_vdms} illustrates a concurrency analysis for
the \textit{1tag} query using VDMS to investigate the number of
concurrent clients to use for the per-query performance analysis.
This figure shows the metadata performance plateaus at 32 clients and
the image performance at 56 clients for a simple query.
To further investigate, Figure~\ref{fig:q_throughput_32} and
Figure~\ref{fig:q_throughput_56} illustrate the throughput performance
for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients,
respectively.
On the 56-client experiment, a 10x difference occurs between VDMS and
MySQL in metadata throughput while 32 clients is similar but not as high.
The overall performance for both MySQL and VDMS is better with 56 clients;
therefore, we used this value for our per-query analysis.

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/q1_latency}
\caption{Query 1 - Latency}
\label{fig:q1_latency}
\end{figure}

\subsection{Videos}

\begin{figure*}[ht!]
\centering
\includegraphics[width=\textwidth]{figures/video_overhead}
\caption{Concurrency (left) and Overhead (right)}
\label{fig:video}
\end{figure*}

VDMS provides full support for video storage and operations.
This includes support for encoding/decoding and trans-coding of mp4, avi, and
mov containers, as well as support for xvid, H.263 and H.264 encoders.
This is supported through the Visual Compute Module that provides an abstraction
layer on top of OpenCV (and ffmpeg directly).
All the operations supported for images in VDMS are also supported at the
video and frame level on the API. On top of that, there are a number of
video-specific operations that are supported, such as the interval operations,
that allow users to retrieve clips at different FPS versions of the video.

All this functionality is provided and integrated with the rest of the
metadata API as part of the comprehensive VDMS interface. This makes it
possible for users to interact with metadata and video in a transactional
manner, enabling users to run queries like: "Retrieve all the videos
where there is a lake with probability higher than X, converting all videos
to H.264 mp4 of size 224x244".
In particular, this functionality was used internally to select a subset
of videos with the right licenses for a video summarization application.

To the best of our knowledge, there is no solution that can provide
all the functionality mentioned above. Also, implementing a baseline is a very
complex task as there is a large number of options and parameters that can
be chosen, which makes it hard to accurately compare against VDMS functionality.
For this reason, we chose to make a study using VDMS in various scenarios,
and analysis what is the impact of having the overhead of VDMS' Request
Server in the overall access time.

Figure \ref{fig:video} shows the analysis of different queries aimed to retrieve
a video using VDMS interface. We show how VDMS increases throughput of serving
a video object as the number of simultaneous client increases, as well as the
overhead VDMS introduces in the overall query execution time.
The figure on the left compares the number of video transaction per second
(i.e., number of videos returned per second) when different operations
are executed as part of the transaction. The upper-bound of this would be
simply returning the video as-is (without running any encoding/decoding or
operation), represented by the red line. This query is the upper-limit because
it essentially translates to reading the video from the file-system and sending
it over a TCP/IP socket, without any other overhead or operations.

We also run a set of other queries that involve: (a) running a resize operation
on the video and, consequently, needs a decoding and
encoding operations as well (blue line),
(b) transcoding, meaning the use a different container and encoder
than the one originally used (yellow line), and (c) both resize and transcoding.
Note that the resize operation performs a downsize, which translates in less
data being sent over the wire. This is specially noticeable when supporting 32
simultaneous clients, where the systems provides more videos per second due to
sending less data to the client, when compares to just transcoding and not resizing (yellow line).

We can see that the system performs best when using all the physical cores.
COMPLETE THIS BETTER.

WE SHOULD ADD HERE SAMPLES OF THE QUERYS TO SHOW HOW THESE ARE RETRIEVED.

Because we see almost 3 orders of magnitude drop in performance when including
operations as part of the query, we wanted to understand where most of the time
was spent on the query, and optimize the Request Server and Visual Compute Module
if necessary. For this, we run the experiment shown at
Figure \ref{fig:video} (right) which breaks down the different components of the
queries. This figure shows that more than 97\% of the query execution is spent
on encoding/decoding operations, which is well-known to be a
heavy and compute intensive operation.
On the one hand, this results show that VDMS introduced overhead for
video operation is minimal. On the other hand, this result means a
limit on the opportunities for optimization for video queries given
that biggest time factors are accounted by encoding/decoding, which is
outside the scope of VDMS.
This result also was the inspiration point for one optimization we included
in future versions of VDMS, which involves using ffmpeg C++ API to
limit the number of frames being encoding/decoding when possible.

\subsection{Feature Vectors}

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities or visual objects.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
These vectors can be labeled and classified to build search indexes,
and there are many in-memory libraries that are designed for
this task~\cite{flann, faiss}.
Using the VDMS API, users can manage feature vector indexes,
query previously inserted elements (images),
run a k-nearest neighbor search (knn) and express relationships
between existing images or feature vectors and
the newly inserted feature vectors.
By natively supporting feature vectors and knn,
VDMS allows out-of-the-box classification
functionalities for many applications.

For this work, we wanted to

% Figure~\ref{fig:features} shows an example of how this functionality
% can be integrated in a medical imaging analytics pipeline:
% (1) A feature vector is extracted from an bounding
% box in a brain image and labeled;
% (2) the feature vector is inserted together
% with all the associated metadata (type of cancer, for example)
% and a link to the original image.
% (3) A new feature vector is extracted from a new image,
% (4) a query to VDMS is issued to classify that feature vector based on the
% indexed features in VDMS, and
% (5) VDMS can respond to the query with the label
% associated to that feature vector.
% Note that, even if the example is based on the medical images used for the
% performance evaluation, this methodology is applicable in many other contexts
% and use cases, such as face detection and matching.
% More details about the JSON API for this functionality can be found on
% our Github wiki page.

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/feature_img_results}
\caption{Sample Results of Similarity Search}
\label{fig:similarity}
\end{figure*}

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/features_alternatives}
\caption{Feature Vector Evaluation}
\label{fig:features_eval}
\end{figure*}

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/features_disksize}
\caption{Feature Collection Size in Disk}
\label{fig:features_size_does_matter}
\end{figure*}