### 2020-02-17 yfcc100m/paper/vldb2020/intro.tex

\section{Introduction}
\label{intro}

Visual computing workloads performing analytics on
video or image data, either off-line or streaming,
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning techniques to
extract information out of the visual data which can subsequently be used
for informed decision making and analytics\cite{vdms-nips}.
Visual data can provide important insights for various applications:
a retail vendor might be interested in the amount of time
shoppers spend in front of a specific product, a medical expert might
want to study the effect of a specific treatment on the size of a tumor by
looking at brain scans, and a media company would like to leverage
thousands of hours of produced video material for different uses.

Moreover, there are many applications that can be studied through
the use of large and public available datasets.
But before getting started, there is usually
a large effort devoted to cleaning up the data.
For this work, we use the YFCC100m dataset because of its comprehensive nature,
and mainly because we have used it for multiple proof of concepts
and applications within our research lab at Intel.
Applications include basic image search functionality (through the use
of human-generated tags), advanced image search through the use of
machine generated tags for each image and feature vectors,
and video summarization.
In all cases, it was needed to select a subset of the data before running
any processing. Selecting subsets of data is by itself a time consuming task,
as it involves loading all metadata into a solution that enables searches
based on properties (some of the common tools used for this task are
relational database, graph database, csv files, etc), and
building the necessary pipelines and plug-ins
for querying and retrieving the right data.

Despite the fact that visual data is becoming a primary component in analytics,
there has been very little research on the management of visual data, even
less that resulted in fully functioning and available tools.
Most of the current storage solutions are an ad-hoc collection of
tools combined with custom scripts to tie them together, unique not only to
a specific discipline but often to individual researchers.
For example, consider an ML developer constructing a pipeline
for extracting brain tumor information from existing brain images in a
classic medical imaging use case. This task requires assigning consistent
identifiers for the scans and adding their metadata in
some form of relational or key-value database.
If the queries require search over some patient information,
then patients have to be associated with their brain scans images.
Finally, if the ML pipeline needs images that are of a size different
than the stored ones, there is additional compute diverted
towards preprocessing after the potentially larger images are fetched.
All these steps require investigation of different software
solutions that provide various functionalities that have to be adapted,
re-purposed, and stitched together with scripts for each specific use case.
Moreover, most databases make it hard to evolve the schema on the fly and
require a definition of the representation of the data beforehand, which becomes
a problem when new forms of data becomes available with the advances on ML based
data extraction techniques.
Not only do these ad hoc solutions make replicating experiments
difficult, they do not scale well to real-world applications.
Addressing the storage and retrieval of visual data necessitates a complete
overhaul of the storage architecture, preferably using emerging breakthroughs in
heterogeneous memory and storage for efficiency, and with a focus to recognizing
types of data that becoming the center of analytics (i.e., images and videos).

In this paper, we show how a new approach for visual data management can
a very complex task become simple and productive.
We also show how, buy using a tool that was design for the job,
inefficiencies that emerge from the re-purpose of multiple system disappear
and there are important gains in overall data access performance.
We put a particular emphasis on the scalability and concurrenct aspect of the
performance evaluation given that large datasets are processed in parallel
using distributed computing.

We also expand on the Video and Feature Vector capabilities of
VDMS, which are part of the latest additions to the system.
We analyze different functionalities and trade-offs for this type of data,
in combination with metadata filtering.
To the best of our knowledge, this set of functionalities are unique to
VDMS and we were not able to find a system with similar functionality.

We show how VDMS can be used as the single and centralize point for data
management and data access even when having multiple modalities of data:
Metadata, Image, Videos, and Feature Vectors, all part of a comprehensive
dataset that is used for multiple use-cases.

We start by briefly describing VDMS design principles and implementation
in Section \ref{arch}.
We then describe the dataset we used for our experiments and explain
why this is a comprehensive use-case for visual data analytics, as well as
describe our evaluation methodology in Section \ref{eval}.
Finally, we show the results of our evaluation in Section \ref{eval}.