### 2020-02-17 yfcc100m/paper/vldb2020/legal.tex

\scriptsize
\begin{framed}

2020 Intel Corporation.

Intel and Intel Core are trademarks of Intel Corporation in
the U.S. and/or other countries. Other names and brands may be claimed as the property of others.

Performance results are based on testing as of 10/02/2020 and may not reflect
all publicly available security updates.
See configuration disclosure for details.
No product can be absolutely secure.

Software and workloads used in performance tests may have been optimized for
performance only on Intel microprocessors. Performance tests, such as SYSmark
and MobileMark, are measured using specific computer systems, components,
software, operations and functions.
Any change to any of those factors may cause the results to vary.
You should consult other information and performance tests
to assist you in fully evaluating your contemplated purchases, including the
performance of that product when combined with other products.
For more information go to http://www.intel.com/performance.

Intel's compilers may or may not optimize to the same degree for non-Intel
microprocessors for optimizations that are not unique to Intel microprocessors.
These optimizations include SSE2, SSE3, and SSSE3 instruction
sets and other optimizations. Intel does not guarantee the availability,
functionality, or effectiveness of any optimization on microprocessors not
manufactured by Intel. Microprocessor-dependent optimizations in this product
are intended for use with Intel microprocessors. Certain optimizations not
specific to Intel microarchitecture are reserved for Intel microprocessors.
Please refer to the applicable product User and Reference Guides for more
information regarding the specific instruction sets covered by this notice.
\end{framed}