### 2020-02-25 yfcc100m/paper/vldb2020/acks.tex
Update to overleaf version

\section{Acknowledgments}

We would like to thank the many people that made this project possible and
helped us through the process, as this work is the results of many efforts.
We want to specially thank our senior technologists Nilesh Jain and Ravi Iyer for their
full support and input during the duration of the project.
We want to specially acknowledge Philip Lantz and Vishakha Gupta for their help
with PMGD, key to load large datasets into VDMS.
We want to thank Jim Blakley for his input during the various 
phases of our project, and for advocating and promoting its use.
We want to acknowledge the Intel Labs VDMS team for their efforts
open-sourcing and maintaining the system.
We want to thank Jason Gardner for his helping in setting up many of the servers
and infrastructure needed to conduct our experiments.