### 2020-02-25 yfcc100m/paper/vldb2020/conclusion.tex

\section{Conclusion}

In this paper, we show a comprehensive evaluation of VDMS 
using one of the largest publicly available datasets: 
The Yahoo Flickr Creative Commons 100M (YFCC100M), 
together with the expansions packs that include 
machine-generated labels and feature vectors.
We compare VDMS against a combination of industry standard systems, 
all of which are needed to replicate a portion of VDMS' 
functionality in order to implement an Image Search application. 
We see improvements up to 35x in certain queries, 
and an average improvement of about 15x. 
VDMS also supports video and 
we provide an analysis of the performance of the video querying
on a highly concurrent setting. 
Finally, we analyze the different trade-offs of VDMS' 
descriptors indexes, a functionality 
that fully integrates with the rest of VDMS interface, 
providing a powerful and comprehensive API.
The design of VDMS, which was conceived as a 
data management system that treats visual entities 
as first class citizens, can remove inefficiencies 
that result from re-purposing and combining solutions
that were not designed for the job while providing 
simpler and richer interfaces.