### 2020-02-25 yfcc100m/paper/vldb2020/intro.tex

\section{Introduction}
\label{intro}

Visual computing workloads performing analytics on
video or image data, either off-line or streaming,
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning (ML) techniques to
extract information from the visual data which can subsequently be used
for informed decision making \cite{vdms-nips}.
The insights this information can provide depend on the
application: a retail vendor might be interested in the amount of time
want to see the effect of a specific treatment on the size of a tumor.

Despite this rich and varied usage environment, there has been very little
research on the management of visual data.
Most of the current storage solutions for visual data are
an ad-hoc collection of tools and systems. 
For example, consider a ML developer constructing a pipeline
for extracting brain tumor information from existing brain images in a
classic medical imaging use case. 
This requires assigning consistent
identifiers for the scans and adding their metadata in
some form of relational or key-value database. 
If the queries require a search over some patient information, 
then patients have to be associated with their brain scans. 
Finally, if the ML pipeline needs images that
are of a size different than the stored ones, there is additional compute
diverted towards pre-processing after the potentially larger images are fetched. 
All these steps require investigation of different software
solutions that provide various functionalities that can then be stitched
together with a script for this specific use case.
Moreover, if the pipeline identifies
new metadata to be added for the tumor images, most databases make it
hard to evolve the schema on the fly.
As another example, many applications that can be studied through the use of large
and publicly available datasets. 
Applications include basic image search functionality (through the use
of human-generated tags), advanced image search through the use of
machine-generated tags and feature vectors\cite{imagesearch} 
for each image, and video summarization.
For these use-cases, the usual first step consists on selecting a 
subset of the data before running any processing, and a large effort 
is devoted to cleaning up and pre-processing the data.
Selecting subsets of data is by itself a time consuming task,
as it involves loading all metadata into a solution that enables searching
based on tags (relational database, graph database, csv files, etc), and
building the necessary pipelines for querying and retrieving the right data.

More generally, data scientists and machine learning developers 
usually end up building an ad-hoc solution that results in a 
combination of databases and file systems to store 
metadata and visual data (images, videos), respectively. 
This is integrated with a set of custom scripts that tie multiple systems together, 
unique not only to a specific application/discipline but often to individual researchers.
These ad-hoc solutions make replicating experiments difficult, 
and more importantly, they do not scale well when deployed in the real-world.
The reason behind such complexity is the lack of a one-system 
that can be used to store and access all the data the application needs.

In this paper, we show how VDMS~\cite{vdms-nips} provides a comprehensive solutions 
to the data management for applications that heavily rely on visual data. 
VDMS is an Open Source project designed to enable efficient access of visual data.
We also expand on the video and feature vector capabilities of
VDMS, which are part of the latest additions to the system.
We analyze different functionalities and trade-offs for this type of data,
in combination with metadata filtering. 
To the best of our knowledge, this set of functionalities, 
provided behind an integrated API, are unique to VDMS and 
we were unable to find a system with similar functionality.
We show how VDMS can be used as the single and centralize point for data
management and data access even when having multiple modalities of data:
Metadata, Image, Videos, and Feature Vectors.

For this work, we use the YFCC100M dataset\cite{Thomee_2016}. 
The YFCC100M is the largest publicly multimedia collection. 
It contains the metadata of around 99.2 million photos 
and 0.8 million videos from Flickr,
plus expansion packs that include a variety of multidimensional data,
all of which were shared under one of the various Creative Commons licenses.
We have used this dataset
for multiple proof of concepts and applications within our research lab.