### 2020-02-27 yfcc100m/paper/vldb2020/acks.tex
Paper submitted Feb 26

\section{Acknowledgments}

We would like to thank the many people that made this project possible and
helped us through the process, as this work is the results of many efforts.
We want to specially thank our senior technologists Nilesh Jain and 
Ravi Iyer for their full support and input during the duration of the project.
We want to specially acknowledge Philip Lantz and Vishakha Gupta for their help
with PMGD, key to loading large datasets into VDMS.
We want to thank Jim Blakley for his input during the various 
phases of our project, and for advocating and promoting VDMS.
We want to acknowledge the Intel Labs VDMS team for their efforts
open-sourcing and maintaining the system.
We want to thank Jason Gardner for his helping in setting up many of 
the servers and infrastructure needed to conduct our experiments.