### 2021-01-20 yfcc100m/paper/vldb2021/intro.tex

\section{Introduction}
\label{intro}

Visual computing workloads performing analytics on videos and/or images
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning (ML) techniques to
extract information from the visual data which can subsequently be used
for informed decision making \cite{vdms-nips}.
The insights this information can provide depend on the
application: retail vendors might be interested in knowing 
which are the most visited areas of their stores using security 
video feeds as input, or a doctor might want know the effect of a
specific treatment by looking at the changes in size of a tumor from a brain scan.

Despite the increasing use of visual data processing, 
there has been very little research on the \textit{management} of visual data.
Most of the current storage solutions for visual data are
an ad-hoc collection of tools and systems, that are re-purposed and 
adapted to work with visual data.
The approach of re-purposing and integrating solutions not designed for a task
results in resource utilization inefficiencies\cite{rasdaman}. 
For example, consider a ML developer constructing a pipeline
for extracting brain tumor information from existing brain images in a
classic medical imaging use case. 
This requires assigning consistent
identifiers for the scans and adding their metadata in
a relational or key-value database. 
If the queries require a search over patient information, 
then patients are associated with their brain scans. 
Finally, if the ML pipeline needs images with a different resolution than the original, there is additional compute
diverted towards pre-processing the original images which are typically larger. 
All these steps require understanding different software
solutions that provide various functionalities that can then be stitched
together with a script for this specific use case.
Moreover, if the pipeline identifies
new metadata to be added for the tumor images, most databases make it
hard to change the schema on the fly.
As another example, many applications can be studied through the use of large
and publicly available datasets. 
Applications include basic image search functionality (through the use
of human-generated tags), advanced image search through the use of
machine-generated tags and feature vectors\cite{imagesearch} 
for each image, and video summarization.
For these use-cases, the usual first step consists on selecting a 
subset of the data before running any processing, and a large effort 
is devoted to cleaning and pre-processing the data.
Selecting subsets of data is by itself a time consuming task,
as it involves loading all metadata into a solution that enables searching
based on tags (relational database, graph database, csv files, etc), and
building the necessary pipelines for querying and retrieving the right data.

More generally, data scientists and machine learning developers 
usually end up building an ad-hoc solution that results in a 
combination of databases and file systems to store 
metadata and visual data (images, videos), respectively. 
This is integrated with a set of custom scripts that tie multiple systems together, 
unique not only to a specific application/discipline but often 
to individual researchers or development teams.
These ad-hoc solutions make replicating experiments difficult, 
and more importantly, they do not scale well when deployed in the real-world.
The reason behind such complexity is \textit{the lack of a system 
that can be used to store and access all the data the application needs, 
including metadata, images, videos, and feature vectors}.

In this paper, we describe VDMS and show how it provides a 
comprehensive solution to the data management for applications 
that heavily rely on visual data. 
VDMS is a completely Open Source project designed to enable 
efficient access of visual data.
To the best of our knowledge, a rich set of functionalities 
designed for visual data management, provided behind an integrated API, 
is unique to VDMS and we were unable to find a system with similar functionality.
While there are a number of big-data frameworks~\cite{spark, hadoop}, systems
that can be used to store metadata~\cite{memsql, vertica}, and systems that
manipulate a specific category of visual data~\cite{scidb, rasdaman}, VDMS can
be distinguished from them on the following aspects:

\begin{itemize}
\item {\em Design for analytics and machine learning}: By targeting
visual data for use cases that require manipulation
of visual information and associated metadata,
\item {\em Ease-of-use}: By defining a common API that allows applications to
combine their complex metadata searches with operations on resulting visual
data, and together with full support for feature vectors. VDMS goes beyond the
traditional SQL or OpenCV level interfaces that do one or the other.
\item {\em Performance}: We show how a unified system such as VDMS can
outperform an ad-hoc system constructed with well-known discrete components.
Because of the capabilities we have built into VDMS, it handles complex
queries significantly better than the ad-hoc system without compromising the
performance of simple queries.
\end{itemize}

In order to evaluate VDMS in a realistic use case, 
we use the YFCC100M dataset\cite{Thomee_2016}. 
The YFCC100M is the largest public multimedia collection. 
It contains the metadata of around 99.2 million photos 
and 0.8 million videos from Flickr,
plus expansion packs that include a variety of multidimensional data,
all of which were shared under one of the various Creative Commons licenses.
We have used this dataset
for multiple proof of concepts and applications within our research lab.

We show how VDMS can be used as a centralize point for data
management and data access even when having multiple modalities of data:
Metadata, Image, Videos, and Feature Vectors.