### 2021-04-22 yfcc100m/paper/vldb2021/reviewer_response.txt
added file for reviewer responses and fixed typos from reviewer 5

Reviews: Point addressed
========================================
Reviewer #1

Reviewer #4

Reviewer #5
    D4: Fixed all typos