### 2021-04-28 yfcc100m/paper/vldb2021/futurework.tex
Add future work section

\section{Future Work}

Add future work here, responding reviewrs questions.