### 2021-04-28 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex

\documentclass[11pt]{proposalnsf}
\usepackage{times}
\usepackage{mathptmx}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{array}
\usepackage[noend,nofillcomment]{algorithm2e}
\usepackage{verbatim}
%\usepackage{algorithmic}
%\usepackage{amsfonts}
%\usepackage{lineno}
\usepackage{ifthen}
\usepackage{booktabs}

\usepackage{color}

\usepackage{multirow}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)

\makeatletter
\def\tightitemize{\ifnum \@itemdepth >3 \@toodeep\else \advance\@itemdepth \@ne
\edef\@itemitem{labelitem\romannumeral\the\@itemdepth}%
\list{\csname\@itemitem\endcsname}{\setlength{\topsep}{-\parskip}\setlength{\parsep}{0in}\setlength{\itemsep}{0in}\setlength{\parskip}{0in}\def\makelabel##1{\hss\llap{##1}}}\fi}
\let\endtightitemize =\endlist
\def\tightenumerate{%
\ifnum \@enumdepth >\thr@@\@toodeep\else
\advance\@enumdepth\@ne
\edef\@enumctr{enum\romannumeral\the\@enumdepth}%
\expandafter
\list
\csname label\@enumctr\endcsname
{\setlength{\topsep}{-\parskip}\setlength{\parsep}{0in}\setlength{\itemsep}{0in}\setlength{\parskip}{0in}\usecounter\@enumctr\def\makelabel##1{\hss\llap{##1}}}%
\fi}
\let\endtightenumerate =\endlist
\makeatother
\renewcommand{\refname}{\centerline{References cited}}

% this handles hanging indents for publications
\def\rrr#1\\{\par
\medskip\hbox{\vbox{\parindent=2em\hsize=6.12in
\hangindent=4em\hangafter=1#1}}}

\def\baselinestretch{1}

\begin{document}

\begin{center}
{\Large{\bf VLDB 2021 - Notes for Reviewers}}

\end{center}

Here we address rebviews comments 1 by one.

\newpage
\pagenumbering{arabic}

\end{document}