### 2021-04-28 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex
Update response doc

\documentclass[11pt]{proposalnsf}
\usepackage{times}
\usepackage{mathptmx}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{array}
\usepackage[noend,nofillcomment]{algorithm2e}
\usepackage{verbatim}
%\usepackage{algorithmic}
%\usepackage{amsfonts}
%\usepackage{lineno}
\usepackage{ifthen}
\usepackage{booktabs}

\usepackage{color}

\usepackage{multirow}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)

\makeatletter
\def\tightitemize{\ifnum \@itemdepth >3 \@toodeep\else \advance\@itemdepth \@ne
\edef\@itemitem{labelitem\romannumeral\the\@itemdepth}%
\list{\csname\@itemitem\endcsname}{\setlength{\topsep}{-\parskip}\setlength{\parsep}{0in}\setlength{\itemsep}{0in}\setlength{\parskip}{0in}\def\makelabel##1{\hss\llap{##1}}}\fi}
\let\endtightitemize =\endlist
\def\tightenumerate{%
\ifnum \@enumdepth >\thr@@\@toodeep\else
\advance\@enumdepth\@ne
\edef\@enumctr{enum\romannumeral\the\@enumdepth}%
\expandafter
\list
\csname label\@enumctr\endcsname
{\setlength{\topsep}{-\parskip}\setlength{\parsep}{0in}\setlength{\itemsep}{0in}\setlength{\parskip}{0in}\usecounter\@enumctr\def\makelabel##1{\hss\llap{##1}}}%
\fi}
\let\endtightenumerate =\endlist
\makeatother
\renewcommand{\refname}{\centerline{References cited}}

% this handles hanging indents for publications
\def\rrr#1\\{\par
\medskip\hbox{\vbox{\parindent=2em\hsize=6.12in
\hangindent=4em\hangafter=1#1}}}

\def\baselinestretch{1}

\begin{document}

\begin{center}
{\Large{\bf VLDB 2021 - Response to Reviews: \\
Using VDMS to Index and Search 100M Images}}

\end{center}

\section{Reviewer \#1}

\textbf{
D1 (W1): Although VDMS mentions that it supports Visual Feature Vectors as
first class citizens in Section 3.3, the queries used in evaluation
do not include an example.
}\bigskip

Not for this work, we will add to future work.
Also, building a comparison baseline that integrates all that functionality
is very ad-hoc, and would complicate the understanding of the origins
of the benefits of VDMS.

We can provide appendix.

\bigskip
\textbf{
D2 (W1): All queries involve “resize” operation, which is expensive for the
baselines since it is happening on the client side.
A simplistic query without resizing can show how the graph model performs
against traditional RDS DBs (i.e., will show that VDMS has comparable
performance even for such use-case without "visual manipulation").
}\bigskip

This is a great point, we appreciate the suggestion.
We have added

\bigskip
\textbf{
D3 (W2): It is not clear which queries are supported or not.
The queries used for evaluation are based on tags and geo search.
What about visual-based queries? For example, is image similarity search based
on Visual Feature Vectors within a geographical region supported?
}\bigskip

VDMS do support the suggested query, i.e., perform a similarity search based
on feature vectors and also filter by geographical location.
This can be achieved using the very similar queries as the one used in
the present evaluation, plus the use of the feature vector queries,
that are natively supported using the same set of APIs.
Because of space constraints, we were not able to accommodate initial evaluation
we have been working on with Feature Vector search specifically,
and we plan to perform a more comprehensive evaluation on this functionality
of this feature because of its relevance when processing visual data, as well
as our own needs in our use cases.

\bigskip
\textbf{
D4 (W2): Section 4.3.1, paragraph 5, mentions that INTERSECTION operation is not
yet implemented at the server side.
A discussion on current limitations/future work for such queries or queries
such as in D3, will be better to be clarified in paper.
}\bigskip

Thanks for this comment, we have added a "Future Work" section where we cover
this and other question, expanding further on the most immediate lines of
improvement and research for the VDMS project.

\bigskip
\textbf{
D5 (W3): Section 4.2 paragraph 1, desing
}\bigskip

Fixed in final version.

\bigskip

\newpage
\section{Reviewer \#4}

\newpage
\section{Reviewer \#5}

\pagenumbering{arabic}

\end{document}