### 2021-04-28 yfcc100m/paper/vldb2021/futurework.tex
Add mention to union/intersetction operations in future work

\section{Future Work}

Based on the results of this evaluation and our experience with VDMS
for this and other internal use cases, we have identified certain features
that are needed to fully exploit the potential of the unified back-end
for visual data infrastructure.
In particular, we have identified the urgent need for a set of UNION/INTERSECTION
interfaces to enable richer queries and improve performance,
given the limitations discussed on \ref{concurrency_analysis}.