### 2021-05-05 yfcc100m/paper/vldb2021/futurework.tex
Update future work

\section{Future Work}

Based on the results of this evaluation and our experience with VDMS
for this and other internal use cases, we have identified certain features
that are needed to fully exploit the potential of the unified back-end
for visual data infrastructure.
In particular, we have identified the urgent need for a set of UNION/INTERSECTION
interfaces to enable richer queries and improve performance,
given the limitations discussed on \ref{concurrency_analysis}.
Even if this evaluation focuses on image search using metadata properties,
VDMS has the capability to also perform similarity search based on
feature vectors.
The team will work on a evaluation of this feature as future work,
as more work is needed to arrive at a comparable and fair baseline.
Last, but not least, we recognize the need for a distributed
implementation of VDMS.
Most of the subcomponents and abstractions were design with a distributed
architecture in mind, but also most of the applications we encountered in our
internal use cases would fit on a single server plus a distributed file system.
With a validation of the benefits of the abstractions provided by VDMS,
the natural next step is working on scaling the system out.